---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  dpi =  300,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geovctrs

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
![R-CMD-check](https://github.com/paleolimbot/geovctrs/workflows/R-CMD-check/badge.svg)
[![Codecov test coverage](https://codecov.io/gh/paleolimbot/geovctrs/branch/master/graph/badge.svg)](https://codecov.io/gh/paleolimbot/geovctrs?branch=master)
<!-- badges: end -->

The goal of geovctrs is to provide a common set of classes and data structures to ensure that processing functions in the rapidly expanding R geospatial ecosystem are interchangeable.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("paleolimbot/geovctrs")
```

You'll have to install GEOS to build this package yourself. There will come a day when this package won't require GEOS, but in the words of [Aragorn](https://en.wikipedia.org/wiki/Aragorn), that day is not this day. On Windows this is taken care of through [rwinlib](https://github.com/rwinlib), on MacOS you can install [homebrew](https://brew.sh/) and run `brew install geos`, and on Linux you can install GEOS from your favourite package manager (`apt-get install libgeos-dev` on Debian/Ubuntu).

If you can load the package, you're good to go!

```{r example}
library(geovctrs)
```

## Geometry vectors

This package provides [vctrs](https://vctrs.r-lib.org/) class definitions for [well-known binary](https://en.wikipedia.org/wiki/Well-known_text_representation_of_geometry#Well-known_binary) (`geo_wkb()`), [well-known text](https://en.wikipedia.org/wiki/Well-known_text_representation_of_geometry) (`geo_wkt()`), and several simple geometries that can be efficiently stored using column vectors (`geo_xy()`, `geo_segment()`, and `geo_rect()`). These classes are designed to work with [dplyr](https://dplyr.tidyverse.org/), [tidyr](https://tidyr.tidyverse.org/), and other [tidyverse](https://tidyverse.org/) packages that use vctrs. In addition to providing default implementations of [ generics](#generics), they have print, plot, and coercion, methods so that they "just work".

```{r}
head(geo_example_wkt)
head(as_geo_wkb(geo_example_wkt))
```

## Constructing and destructing geometries

The geovctrs package provides functions to construct geometries from coordinates, and destruct geometries to extract their coordinates.

```{r}
# construct linestrings
linestrings <- c(
  geo_linestring(geo_xy(c(1, 2, 5), c(0, 1, 2))),
  geo_linestring(geo_xy(c(10, 20, 50), c(0, 10, 20)))
)

linestrings

# destruct to get coordinates
geo_coordinates(linestrings)
```

You can use `separate_xy()` get the actual x and y values (and `unite_xy()` to create a `geo_xy()` column).

```{r}
separate_xy(geo_coordinates(linestrings), "xy")
```

In the [upcoming release of dplyr](https://www.tidyverse.org/blog/2020/03/dplyr-1-0-0-summarise/), this is useful in conjunction with `group_by()` and `summarise()`.

```{r, message=FALSE, warning=FALSE, eval =  FALSE}
library(dplyr)
geo_coordinates(linestrings) %>% 
  group_by(feature) %>% 
  summarise(geometry = geo_linestring(xy))
#> # A tibble: 2 x 2
#>   feature geometry   
#> *   <int> <clctn>    
#> 1       1 / (1 0)…+2 
#> 2       2 / (10 0)…+2
```

## Generics

There are several concepts that show up on repeat in geometry packages. The geovctrs package provides these as generics with reasonable implementations for the bundled [geometry vector classes](#geometry-vectors). Notably, `geo_bbox()` (returns a `geo_rect()`), `geo_plot()` (thin wrapper around `graphics::plot()`), and `geo_summary()` (returns a `tibble()` with basic information about each feature). These generics work on anything that can be interpreted as a geometry vector, including character vectors (interpreted as well-known text), data frames with exactly one geometry column (interpreted as the geometry column), and anything that implements `as_geovctr()` (e.g., [sf](https://r-spatial.github.io/sf) objects).

```{r ex-plot}
geo_bbox(geo_nc)
geo_plot(geo_nc)
```

The geovctrs package also provides a framework for transformers, or functions that accept a vector of geometries and return a vector of geometries. These always return the same type as the input, as dictated by the implementations of `as_geovctr()` and `restore_geovctr()`. This enables transforming functions to work on a wide variety of input types, including sf objects:

```{r}
library(sf)
sf_nc <- read_sf(system.file("shape/nc.shp", package = "sf"))
geo_envelope(sf_nc)
```

See `vignette("extending-geovctrs", package = "geovctrs")` for instructions on how to create a class that works with the geovctrs framework.
