---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  dpi =  300,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geovctrs

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
![R-CMD-check](https://github.com/paleolimbot/geovctrs/workflows/R-CMD-check/badge.svg)
[![Codecov test coverage](https://codecov.io/gh/paleolimbot/geovctrs/branch/master/graph/badge.svg)](https://codecov.io/gh/paleolimbot/geovctrs?branch=master)
<!-- badges: end -->

The goal of geovctrs is to provide a common set of classes and data structures to ensure that processing functions in the rapidly expanding R geospatial ecosystem are interchangeable.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("paleolimbot/geovctrs")
```

If you can load the package, you're good to go!

```{r example}
library(geovctrs)
```

## Geometry vectors

This package provides [vctrs](https://vctrs.r-lib.org/) class definitions several simple geometries that can be efficiently stored using column vectors (`geo_xy()`, `geo_xyz()`, `geo_segment()`, and `geo_rect()`). Well-known geometry (well-known binary and well-known text) vectors from the [wk](https://paleolimbot.github.io/wk) package are used for generic geometries. These classes are designed to work with [dplyr](https://dplyr.tidyverse.org/), [tidyr](https://tidyr.tidyverse.org/), and other [tidyverse](https://tidyverse.org/) packages that use vctrs. In addition to providing default implementations of [generics](#generics), they have print, plot, and coercion methods so that they "just work" when you need them to.

```{r}
head(geo_example_wkt)
head(as_wkb(geo_example_wkt))
```

## Constructing and destructing geometries

The geovctrs package provides functions to construct geometries from coordinates, and destruct geometries to extract their coordinates.

```{r}
# construct linestrings
linestrings <- c(
  geo_linestring(geo_xy(c(1, 2, 5), c(0, 1, 2))),
  geo_linestring(geo_xy(c(10, 20, 50), c(0, 10, 20)))
)

linestrings

# destruct to get coordinates
geo_coordinates(linestrings)
```

You can use `separate_xy()` get the actual x and y values (and `unite_xy()` to create a `geo_xy()` column).

```{r}
separate_xy(geo_coordinates(linestrings), "xy")
```

In the [newest version of dplyr](https://www.tidyverse.org/blog/2020/03/dplyr-1-0-0-summarise/), this is useful in conjunction with `group_by()` and `summarise()`.

```{r, message=FALSE, warning=FALSE}
library(dplyr)
geo_coordinates(linestrings) %>% 
  group_by(feature) %>% 
  summarise(geometry = geo_linestring(xy))
```

## Generics

There are several concepts that show up on repeat in geometry packages. The geovctrs package provides these as generics with reasonable implementations for the bundled [geometry vector classes](#geometry-vectors). Notably, `geo_bbox()`/`geo_envelope()` (return a `geo_rect()`), `geo_srid()`, and `geo_z()`. These generics work on anything that can be interpreted as a geometry vector, including character vectors (interpreted as well-known text), data frames with exactly one geometry column (interpreted as the geometry column), and anything that implements `as_geovctr()` (e.g., [sf](https://r-spatial.github.io/sf) objects).

```{r ex-plot}
geo_bbox(geo_nc)
geo_srid(head(geo_nc))
```

The geovctrs package also provides a framework for transformers, or functions that accept a vector of geometries and return a vector of geometries. These always return the same type as the input, as dictated by the implementations of `as_geovctr()` and `restore_geovctr()`. This enables transforming functions to work on a wide variety of input types, including sf objects:

```{r}
library(sf)
sf_nc <- read_sf(system.file("shape/nc.shp", package = "sf"))
geo_envelope(sf_nc)
```

See `vignette("extending-geovctrs", package = "geovctrs")` for instructions on how to create a class that works with the geovctrs framework.
