---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  dpi =  300,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geovctrs

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
![R-CMD-check](https://github.com/paleolimbot/geovctrs/workflows/R-CMD-check/badge.svg)
[![Codecov test coverage](https://codecov.io/gh/paleolimbot/geovctrs/branch/master/graph/badge.svg)](https://codecov.io/gh/paleolimbot/geovctrs?branch=master)
<!-- badges: end -->

The goal of geovctrs is to provide a common set of classes and data structures to ensure that processing functions in the rapidly expanding R geospatial ecosystem are interchangable.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("paleolimbot/geovctrs")
```

If you can load the package, you're good to go!

```{r example}
library(geovctrs)
```

## Example

This package has reasonable `plot()` and `print()` methods for WKB and WKT types, and provides a [vctrs](https://vctrs.r-lib.org/) implementation for these types so that they "just work" with [tibble](https://tibble.tidyverse.org/), [tidyr](https://tidyr.tidyverse.org/), and [dplyr](https://dplyr.tidyverse.org/) (>=1.0.0) (among others).

```{r}
geo_wkt("LINESTRING (30 10, 10 30, 40 40)")
as_geo_wkb(geo_wkt("LINESTRING (30 10, 10 30, 40 40)"))
tibble(wkt = geo_wkt("LINESTRING (30 10, 10 30, 40 40)"))
```

Most methods work on anything that can be interpreted as geometry, including character vectors and data.frames containing exactly one geometry column.

```{r ex-plot}
geo_plot("LINESTRING (30 10, 10 30, 40 40)")
```

Like any self-respecting geometry package, geovctrs contains a copy of the North Carolina dataset to play with:

```{r}
rev(geo_nc)
```

